### 2022-03-17 omicron-generation.tex
update

For a cohort of individuals who developed symptoms between 1 December 2021 and 2 January 2022, \citep{backer2021omicron} found longer mean (backward) incubation period for the Delta variant than for the Omicron variant (\fref{incubation}A).
However, these measurements were done during a period when the incidence of Omicron was increasing while the increasing of Delta was decreasing (\fref{epidemic}).
Thus, dynamical bias would be expected to lead to shorter observed (backward) incubation periods in Omicron, and longer observed incubation periods in Delta. 
When we account for these growth-rate differences and re-estimate the forward incubation periods, we find that both variants have similar incubation-period distributions (\fref{incubation}B)---for illustrative purposes, we assume $r=-0.05/\mathrm{day}$ and $r=0.15/\mathrm{day}$ for the Delta and Omicron variants, respectively.
Although the exact estimate of the mean forward incubation periods of both variants are sensitive to the assumed growth/decay rates, we find similar means across a plausible ranges of growth rates with unclear differences between two variants (\fref{incubation}C).
% \jd{Do we need the ff.~s.? What does it add to the Fig?}
% \swp{I think people like to see actual values...}
For example, the mean forward incubation period of the Delta variant changes from 3.8 days (95\% CI: 3.5--4.1 days) to 4.4 days (95\% CI: 4.0--4.8 days) as we change the assumed values of $r$ from $-0.1/\mathrm{days}$ to $0/\mathrm{days}$, while
the mean forward incubation period of the Omicron variant changes from 3.8 days (95\% CI: 3.4--4.4 days) to 4.5 days (95\% CI: 3.9--5.5 days) as we change the assumed values of $r$ from $0.1/\mathrm{days}$ to $0.2/\mathrm{days}$.

\begin{figure}[!th]
\includegraphics[width=\textwidth]{figure_incubation.pdf}
\caption{
\textbf{Observed and corrected differences in incubation-period distributions of Delta and Omicron variants.}
(A) Posterior median estimates of the observed (backward) incubation periods of the Delta (black) and Omicron (orange) variants by \cite{backer2021omicron}.
(B) Forward incubation-period distributions assuming $r=-0.05/\mathrm{day}$ and $r=0.15/\mathrm{day}$ for the Delta (black) and Omicron (orange) variants, respectively.
(C) Corrected estimates of the forward incubation-period distributions for different assumptions about the growth rates of the Delta (black, solid lines) and Omicron variants (orange, dashed lines).
Middle lines represent median estimates, and lower and upper lines represent corresponding 95\% confidence intervals.
\label{fig:incubation}
}
\end{figure}

>>>>

For a cohort of individuals who developed symptoms between 1 December 2021 and 2 January 2022, \citep{backer2021omicron} found longer mean (backward) incubation period for the Delta variant than for the Omicron variant (\fref{incubation}A).
However, these measurements were done during a period when the incidence of Omicron was increasing while the increasing of Delta was decreasing (\fref{epidemic}).
Thus, dynamical bias would be expected to lead to shorter observed (backward) incubation periods in Omicron, and longer observed incubation periods in Delta. 
When we account for these growth-rate differences and re-estimate the forward incubation periods, we find that both variants have similar incubation-period distributions (\fref{incubation}B)---for illustrative purposes, we assume $r=-0.05/\mathrm{day}$ and $r=0.15/\mathrm{day}$ for the Delta and Omicron variants, respectively.
Although the exact estimate of the mean forward incubation periods of both variants are sensitive to the assumed growth/decay rates, we find similar means across a plausible ranges of growth rates with unclear differences between two variants (\fref{incubation}C--D).
% \jd{Do we need the ff.~s.? What does it add to the Fig?}
% \swp{I think people like to see actual values...}
For example, the mean forward incubation period of the Delta variant changes from 3.8 days (95\% CI: 3.5--4.1 days) to 4.4 days (95\% CI: 4.0--4.8 days) as we change the assumed values of $r$ from $-0.1/\mathrm{days}$ to $0/\mathrm{days}$ (\fref{incubation}C), while
the mean forward incubation period of the Omicron variant changes from 3.8 days (95\% CI: 3.4--4.4 days) to 4.5 days (95\% CI: 3.9--5.5 days) as we change the assumed values of $r$ from $0.1/\mathrm{days}$ to $0.2/\mathrm{days}$ (\fref{incubation}D).

\begin{figure}[!th]
\includegraphics[width=\textwidth]{figure_incubation.pdf}
\caption{
\textbf{Observed and corrected differences in incubation-period distributions of Delta and Omicron variants.}
(A) Posterior median estimates of the observed (backward) incubation periods of the Delta (black) and Omicron (orange) variants by \cite{backer2021omicron}.
(B) Forward incubation-period distributions assuming $r=-0.05/\mathrm{day}$ and $r=0.15/\mathrm{day}$ for the Delta (black) and Omicron (orange) variants, respectively.
(C--D) Corrected estimates of the mean forward incubation-period for different assumptions about the growth rates of the Delta (C) and Omicron variants (D).
Lines represent median estimates.
Shaded regions represent the corresponding 95\% confidence intervals.
\label{fig:incubation}
}
\end{figure}
